Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-06-28 manuscript/response_to_reviews.Rmd

The study addresses the relationship between drought resistance (Rt) calculated from radial growth, and different traits and microenvironmental characteristics on 12 tree species from one forest in the Eastern US. Both the dataset and the study are meaningful, and the relationship found between Rt and height of biological importance. However, I think there are a number of issues that need to be worked out.

It is important to better discuss the proper use of Rt, particularly when comparing single drought years, to assess tree resistance to drought. This metric has been much used lately as one of the so-called resilience indices (I wonder why other Rs indices where not included in the manuscript). In the paper, the authors use Rt similarly to most studies in the literature, which is fine. Yet, Rt (and particularly the other resilience indices based on radial growth) does not only depend on the individual response to drought events, but also on stand dynamics around target trees. For this reason, if stands around trees are not comparable or show changes after some disturbance (e.g. the drought events analysed) differently, this may bias comparisons of Resilience indices (although it is true that Rt should be less influenced than the other indices). The problem is that there is not such stand competition data for old droughts. This does not invalidate the study but I think the authors should discuss the issue, particularly for the analysis and comparisons of single droughts. I would suggest including also other Resilience indices, not only Rt, to enrich the discussion.

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

**To account for potential trends in individual tree growth prior to the drought, which could be influenced by stand dynamics, we tried an intervention time series analysis (ARIMA). Details on the comparison between $Rt$ and $Rt_{ARIMA}$ are presented in the responses to reviewer 4. In short, the ARIMA method yielded broadly similar results, but tended to produce less reasonable estimates when the difference between the two metrics was large, so we decided to retain $Rt$ as the focal metric. We now include the following content on ARIMA in the manuscript:**

(*insert text*)

@@

**Regarding the resilience indices, we considered analyzing these and decided it would be best to stick with just resistance for three reasons: (1) Most importantly, we would expect resilience to be governed by fundamentally different mechanisms than resistance. While it would be interesting to examine resilience, this would be beyond the scope of the current paper. (2) [DeSoto et al. (2020, Nature)](https://www.nature.com/articles/s41467-020-14300-5) used an extensive dataset of trees sampled after severe droughts and compared the responses to previous, less-severe droughts between trees that survived and trees that were killed by the severe drought. In angiosperms, trees that died from a severe drought had lower resistance to previous droughts, compared to trees that survived the severe drought. In gymnosperms, trees that died during severe droughts had lower recovery following previous droughts, compared to trees that survived the severe drought. These findings would argue that resistance is a more important variable for angiosperms. (3) As the reviewer notes, resilience indices would probably tend to be more sensitive to a tree's neighborhood and prior conditions. We now state the logic for focusing on $Rt$ in the methods section:**

**"We focus exclusively on drought resistance ($Rt$ or $Rt_{ARIMA}$), and not on the resilience metrics described in @lloret_components_2011, because (1) we would expect resilience to be controlled by a different set of mechanisms, and (2) the findings of *[DeSoto et al. (2020)](https://www.nature.com/articles/s41467-020-14300-5)* suggest that $Rt$ is a more important drought response metric for angiosperms. "**

@@

*This is an interesting suggestion. Depending on where we end up going with some other issues, it may make sense to de-emphasize the individual models--perhaps move Table 5 to SI and replace with figure ([issue #71](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71))? However, I think we could also just make an argument for leaving this as is.*

related to [*issue #71*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71)

The authors might like to consider including Table 4 as a supplement, because as it stands it looks like an exploratory analysis. If the ‘single-variable approach’ is not just exploratory, state it clearly and why it is needed besides the full models with several parameters.

[*issue #95*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/95)

@@

L165-167: I think this needs to be explained more clearly. Why are type I errors unlikely and type II error possible? I also think there are real issue with the analysis of crown position. Presumably on average the crown illumination index of trees would have increased over the past 50 or so years (from suppressed to dominant) as trees grew and neighbouring trees died. So I’m not convinced you can include this in your models going back half a century. It’s just not sensible to assume that crown position would not have changed over such a long time period.

L217-226: this is problematic because drought resistance (Rt) as defined by Lloret et al. (2011) assumes that tree growth (BAI in this case) fluctuates around a long term mean that is stable through time. But this is rarely/never the case with tree growth, which shows clear temporal trends related to tree size/ontogeny. If you don’t account for these directional trends then your estimates of Rt will be biased. For instance, if a tree is on a long-term trajectory of decreasing BAI over time, when you calculate Rt you will overestimate the impact of the drought (and vice versa for trees on an upward growth trajectory). So you need to first detrend the tree ring series to remove long-term growth trends (unless you can show that your trees exhibited no long-term temporal trends in BAI). The fact that pre-drought conditions were different for the three droughts is also a problem, but not much you can do about this except acknowledge as you have done.

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

**We do not agree that the $Rt$ metric "assumes that tree growth (BAI in this case) fluctuates around a long term mean that is stable through time." If this was true, $Rt$ would be calculated by dividing growth (BAI) in the drought year by the long-term mean growth increment. Instead, $Rt$ is calculated by dividing growth in the focal year by the mean growth over a narrow window of preceding years (5 years in this study). Even though there may be a long-term directional trend in BAI, if you were to calculate the trend in BAI over each 5-year increment, the trend could go in any direction. To account for such trends, we tried an intervention time series analysis (ARIMA). Details on the comparison between $Rt$ and $Rt_{ARIMA}$ are presented in the responses to reviewer 4. In short, the ARIMA method yielded broadly similar results, but tended to produce less reasonable estimates when the difference between the two metrics was large, so we decided to retain $Rt$ as the focal metric. We now include the following content on ARIMA in the manuscript:**

(*insert text*)

@@

Firstly, the way drought resistance (Rt) is estimated, as the ratio of BAI during drought to the mean BAI over the five years preceding the drought can be problematic, as growth dynamics during this pre-drought period were quite variable across species (Figure 1; incidentally, why do you show ring width indices in this figure if you then use BAI?). This approach has also advantages, particularly when applied in a resilience framework (as in the original Lloret et al. (2011)). However, if you focus specificaly on resistance it seems that directly modelling the growth reduction caused by the drought (e.g., using intervention time series analysis or other methods) would be a more robust approach. At the very least, if you continue using your simpler approach I would strongly suggest including some measure of pre-drought growth in the models to account for differences in growth dynamics among species. It would also nice to clarify why you do not assess recovery and resilience, unlike Lloret et al. (2011).

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

*add specifics (changes to text; SI table/fig #s) to text below*

**We tried an intervention time series analysis (ARIMA). This yielded similar results, which we now present in the Supplementary Info. The original $Rt$ and $Rt_{ARIMA}$ were fairly well correlated, but with some scatter (see figs below). To get a sense of which metric was performing better, we visually reviewed the records of the 18 trees with largest differences between $Rt$ and $Rt_{ARIMA}$. $Rt$ estimates were better for the majority of these, and generally were not hugely unrealistic, whereas $Rt_{ARIMA}$ could be way off (e.g., for several trees that were undergoing growth declines prior to drought and hit a minima at or near the drought, $Rt_{ARIMA}$ was much greater than 1, suggesting a substantial growth increase during the drought.). Thus, we decided to stick with $Rt$ as the primary metric.**

![](tables_figures/Rt_arima_comparison_all_1966.png)
![](tables_figures/Rt_arima_comparison_1977_1999.png)

### 2020-06-28 manuscript/climate_sensitivity_draft.Rmd
added text on ARIMA (needs some details)
#92

For each drought year, we calculated drought resistance ($Rt$) as the ratio of $BAI$ during drought to the mean $BAI$ over the five years preceding the drought [@lloret_components_2011]. Thus, $Rt$ values <1 and >1 indicate growth reductions and increases, respectively. Analyses focused on testing the predictions presented in Table 1, with $Rt$ as the response variable. The general statistical model for hypothesis testing was a mixed effects model with $Rt$ as the response variable, tree nested within species as a random effect, and one or more independent variables as fixed effects. Mixed effects models were implemented in the lme4 package in R [@R-lme4]. We used AICc to assess model selection, and conditional/marginal R-squared to assess model fit, implmented in the AICcmodavg package in R [@R-AICcmodavg].

>>>>

For each drought year, we calculated drought resistance ($Rt$) as the ratio of $BAI$ during drought to the mean $BAI$ over the five years preceding the drought [@lloret_components_2011]. Thus, $Rt$ values <1 and >1 indicate growth reductions and increases, respectively. Because the $Rt$ metric could be biased by directional pre-drought growth trends, we also tried an intervention time series analysis (ARIMA) that predicted drought-year growth based on trends over the past **10** years and used this value in place of the five-year mean in calculations of resistance ($Rt_{ARIMA}$= observed $BAI$/ predicted $BAI$). Because $Rt$ tended to produce more reasonable estimates than $Rt_{ARIMA}$ when there was a large difference between these metrics, we selected $Rt$ as our focal metric, presenting parallel results for $Rt_{ARIMA}$ in the Supplementary Info. We focus exclusively on drought resistance ($Rt$ or $Rt_{ARIMA}$), and not on the resilience metrics described in @lloret_components_2011, because (1) we would expect resilience to be controlled by a different set of mechanisms, and (2) the findings of *[DeSoto et al. (2020)](https://www.nature.com/articles/s41467-020-14300-5)* suggest that $Rt$ is a more important drought response metric for angiosperms.

Analyses focused on testing the predictions presented in Table 1, with $Rt$ as the response variable. The general statistical model for hypothesis testing was a mixed effects model with $Rt$ as the response variable, tree nested within species as a random effect, and one or more independent variables as fixed effects. Mixed effects models were implemented in the lme4 package in R [@R-lme4]. We used AICc to assess model selection, and conditional/marginal R-squared to assess model fit, implmented in the AICcmodavg package in R [@R-AICcmodavg].
